Meta-analysis

### 2020-06-12 04_MetaAnalysis.Rmd
Update meta-analysis and descriptions

   - Grain yield is our response variable.  
   - Trial, which resolves combinations of categorical variables: year, location, row spacing and cultivar; is set as a random intercept.  
   - We will test spray management (our treatment) as a fixed effect and random slope to trial.

```{r LME_slimmer_cluster_disease_mod}
m8 <- lmer(grain_yield.t.ha * 1000 ~ factor(spray_management) +
              (factor(spray_management) | trial_ref),
           data = slimmer_PM_dat)

m9 <- lmer(log(grain_yield.t.ha * 1000) ~ factor(spray_management) +
              (factor(spray_management) | trial_ref),
           data = slimmer_PM_dat)

m10 <- lmer(log(grain_yield.t.ha * 1000) ~
               (factor(spray_management) | trial_ref),
            data = slimmer_PM_dat)

m11 <-
   lmer(log(grain_yield.t.ha * 1000) ~ factor(spray_management) +
           (1 | trial_ref),
        data = slimmer_PM_dat)

anova(m8, m9) # m9 is significantly better model
anova(m9, m10) # m9 is significantly better model
anova(m9, m11) # m11 is a simpler model which is no different to m9
```

The best model from the four above is `m9` with the lower AIC of -29.357.

```{r}
summary(m9)
```

@@

   - A single early spray before first sign of powdery mildew is not likely to increase yields.

   - A single or single spray with one or more follow sprays starting  at the `recommended` first spray , within 3 days of powdery mildew first sign are likely to produce significantly higher grain yields compared to the no spray control.

@@

```{r imputing_variances}
source("./R/impute_yield_err.R")

@@

#### Imputing sample variances from Mean squares
```{r imputing_MSE_dataSpread}
slimmer_PM_dat <- read.csv("cache/slimmer_PM_clusterdat.csv")
hist_usq(unique(slimmer_PM_dat$Y_Msquare))
hist_usq(log(unique(slimmer_PM_dat$Y_Msquare))) # log mean square has a more normal distribution

@@

```{r imputing_MSE}
TrialMSQ <- slimmer_PM_dat %>%
   group_by(trial_ref,location,year)%>%
   summarise(unique(Y_Msquare))

@@

```{r treatment_means_plot}
# class spray_management as a factor and reorder them for the plot
slimmer_PM_dat$spray_management <-
   factor(slimmer_PM_dat$spray_management,
          levels(factor(slimmer_PM_dat$spray_management))[rev(c(1, 2, 4, 5, 3))])

@@

```{r plot_yieldProportion}
slimmer_PM_dat$spray_management <-
   factor(slimmer_PM_dat$spray_management, rev(
      c(
         "control",
         "Early",
         "Recommended",
         "Recommended_plus",
         "Late_plus"
      )
   ))

slimmer_PM_dat %>%
   ggplot(aes(y = prop_yield_gain, x = spray_management)) +
   geom_boxplot() +
   #geom_point(position = "jitter", alpha = 1/5)+
   geom_jitter(width = 0.1, alpha = 1 / 5) +
   labs(x = "Spray management variable",
        y = "Grain yield (t/Ha)",
        title = "Mean grain yield difference to the control for each treatment \n categorised by spray management scenario") +
   theme(plot.title = element_text(hjust = 0.5)) +
   geom_hline(yintercept = 0, linetype = 2) +
   coord_flip()
```

@@

```{r metafor_organisation}
slimmer_PM_dat$spray_management <-
   factor(
      slimmer_PM_dat$spray_management,
      c(
         "control",
         "Early",
         "Recommended",
         "Recommended_plus",
         "Late_plus"
      )
   )

slimmer_PM_dat$yi <- log(slimmer_PM_dat$grain_yield.t.ha)
slimmer_PM_dat$vi <-
   slimmer_PM_dat$Y_Msquare / (slimmer_PM_dat$n * slimmer_PM_dat$grain_yield.t.ha ^
                                  2)

slimmer_PM_dat$spray_management <-
   factor(slimmer_PM_dat$spray_management)
slimmer_PM_dat$trial_ref <- factor(slimmer_PM_dat$trial_ref)
slimmer_PM_dat$trial <- factor(slimmer_PM_dat$trial)
```

```{r Metafor-analysis}
PM_mv_AI <- rma.mv(
   yi,
   vi,
   mods = ~ spray_management,
   method = "ML",
   random = list(~ spray_management | trial),
   struct = "UN",
   data = slimmer_PM_dat
)
```

@@

```{r run_iterations_asJobscript, eval=FALSE, include=FALSE}
# This chunk iterates the imputation of MSEs and reruns the metafor analysis
# each time it iterates it appends some of the output information into the "metaIteration.csv" file
# To increase the speed the calculations/iterations it is send to multiple jobs in the background.
#
# WARNING don't run a high number of iterations without increasing sys.sleep
# sys.sleep should not be reduced below 4 for 50 iterations and
# below 0.5 for 10 iterations
for (ITERATE in 1:10) {
   rstudioapi::jobRunScript(
      "./R/ReRun_metafor-grainyield.R",
      workingDir = here::here(),
      name = "metafor_iteration",
   )
   Sys.sleep(1)
}
```

>>>>

   - Grain yield is our response variable and will be converted to Kg / Ha.  
   - Trial, which resolves combinations of categorical variables: year, location, row spacing and cultivar; is set as a random intercept.  
   - We will test spray management (our treatment) as a fixed effect and random slope to trial.

First lets compare models with and without a log transformed response variable to evaluate if the log transformed model better fits our assumptions of normality.  
Next we will test whether including `spray_management` as a significantly accounts for enough variation as a fixed effect variable to be included in the model.  
Finally we will test whether including `spray_management` as a significantly accounts for enough variation as a random slope to the trial variable.

```{r LME_slimmer_cluster_disease_mod}
m8 <- lmer(grain_yield.t.ha * 1000 ~ factor(spray_management) +
              (factor(spray_management) | trial),
           data = slimmer_PM_dat)

m9 <- lmer(log(grain_yield.t.ha * 1000) ~ factor(spray_management) +
              (factor(spray_management) | trial),
           data = slimmer_PM_dat)

m10 <- lmer(log(grain_yield.t.ha * 1000) ~
               (factor(spray_management) | trial),
            data = slimmer_PM_dat)

m11 <-
   lmer(log(grain_yield.t.ha * 1000) ~ factor(spray_management) +
           (1 | trial),
        data = slimmer_PM_dat)
```

```{r m8_vs_m9}
anova(m8, m9) # m9 is significantly better model
```
There is a significant difference between `m8` and `m9`, with `m9` showing the best fit and lowest AIC: `r min(anova(m8, m9)$AIC)`.

```{r m9_vs_m10}
anova(m9, m10) # m9 is significantly better model
```
There is a significant difference between `m10` and `m9`, with `m9` showing the best fit and lowest AIC: `r min(anova(m9, m10)$AIC)`.

```{r m9_vs_m11}
anova(m9, m11) # m11 is a simpler model which is no different to m9
```

The best model from the four above is `m11` with the lower AIC of -`r min(anova(m9, m11)$AIC)`

```{r}
summary(m11)
```

@@

   - A single `early` spray before first sign of powdery mildew is not likely to increase yields.

   - One spray at the `recommended` timing, or two sprays starting at the `recommended` first spray, are likely to produce significantly higher grain yields compared to the no spray control.

@@

```{r imputing_variances, eval=FALSE, include=FALSE}
source("./R/impute_yield_err.R")

@@

#### Imputing sample variances from Mean squares
```{r imputing_MSE_dataSpread, eval=FALSE, include=FALSE}
slimmer_PM_dat <- read.csv("cache/slimmer_PM_clusterdat.csv")
hist_usq(unique(slimmer_PM_dat$Y_Msquare))
hist_usq(log(unique(slimmer_PM_dat$Y_Msquare))) # log mean square has a more normal distribution

@@

```{r imputing_MSE, eval=FALSE, include=FALSE}
TrialMSQ <- slimmer_PM_dat %>%
   group_by(trial_ref,location,year)%>%
   summarise(unique(Y_Msquare))

@@

```{r treatment_means_plot}
# class spray_management as a factor and reorder them for the plot
slimmer_PM_dat$spray_management <-
   factor(slimmer_PM_dat$spray_management,
          levels(factor(slimmer_PM_dat$spray_management))[rev(c(1, 2, 5, 6, 3,4))])

@@

```{r plot_yieldProportion}
slimmer_PM_dat$spray_management <-
   factor(slimmer_PM_dat$spray_management, rev(
      c(
         "control",
         "Early",
         "Recommended",
         "Recommended_plus",
         "Late",
         "Late_plus"
      )
   ))

slimmer_PM_dat %>%
   ggplot(aes(y = prop_yield_gain, x = spray_management)) +
   geom_boxplot() +
   geom_jitter(width = 0.1, alpha = 1 / 5) +
   labs(x = "Spray management variable",
        y = "Grain yield (t/Ha)",
        title = "Mean grain yield difference to the control for each treatment \n categorised by spray management scenario") +
   theme(plot.title = element_text(hjust = 0.5)) +
   geom_hline(yintercept = 0, linetype = 2) +
   coord_flip()
```

@@

```{r metafor_organisation}
slimmer_PM_dat$spray_management <-
   factor(
      slimmer_PM_dat$spray_management,
      c(
         "control",
         "Early",
         "Recommended",
         "Recommended_plus",
         "Late",
         "Late_plus"
      )
   )

slimmer_PM_dat$yi <- log(slimmer_PM_dat$grain_yield.t.ha)

slimmer_PM_dat$vi <-
   slimmer_PM_dat$Y_Msquare / (slimmer_PM_dat$n * slimmer_PM_dat$grain_yield.t.ha ^
                                  2)

slimmer_PM_dat$trial_ref <- factor(slimmer_PM_dat$trial_ref)
slimmer_PM_dat$trial <- factor(slimmer_PM_dat$trial)
```

```{r Metafor-analysis}
PM_mv_AI <- rma.mv(
   yi,
   yield_error,
   mods = ~ spray_management,
   method = "ML",
   random = list(~ spray_management | trial),
   struct = "UN",
   data = slimmer_PM_dat
)

@@

```{r run_iterations_asJobscript, eval=FALSE, include=FALSE}
# This chunk iterates the imputation of MSEs and reruns the metafor analysis
# each time it iterates it appends some of the output information into the "metaIteration.csv" file
# To increase the speed the calculations/iterations it is send to multiple jobs in the background.
#
# WARNING don't run a high number of iterations without increasing sys.sleep
# sys.sleep should not be reduced below 4 for 50 iterations and
# below 0.5 for 10 iterations
for (ITERATE in 1:50) {
   rstudioapi::jobRunScript(
      "./R/ReRun_metafor-grainyield.R",
      workingDir = here::here(),
      name = "metafor_iteration",
   )
   Sys.sleep(15)
}
```
